Static contract checking for Haskell

### 2011-08-23 paper.tex
Now using Simon's nice ruleform for translations

\newcommand{\etrans}[1]{\mathcal{E} \llbracket #1 \rrbracket}
\newcommand{\dtrans}[1]{\mathcal{D} \llbracket #1 \rrbracket}
\newcommand{\ktrans}[1]{\mathcal{K} \llbracket #1 \rrbracket}
\newcommand{\ctrans}[1]{\mathcal{C} \llbracket #1 \rrbracket}
\newcommand{\trans}[1]{\llbracket #1 \rrbracket}

@@

\begin{figure}[h]
  \centering
  \[
  \begin{array}{rclr}
    mod &:=& def_1,\dots,def_n &\\
    def &\in& \mbox{Definition} & \\
    def &:=& \texttt{data } T = K_1 \mid \cdots \mid K_n & \\
    &\mid& f \in c & \\
    &\mid& f~\vec{x}~=~e & \\
    &\mid& \hspace{-0.2cm}
       \begin{array}{rcl}
         f~\vec{x} &=& \texttt{case } e \texttt{ of } \\
         && K_1(\vec{x}) \to e_1 \mid \cdots \mid K_n(\vec{x}) \to e_n \\
       \end{array} & \\
  \end{array} \]

@@

\begin{figure}
\begin{eqnarray*}
  \etrans{x} &=& x\\
  \etrans{f(e_1,\dots,e_n)} &=& f(\etrans{e_1},\dots,\etrans{e_n})\\
  \etrans{K(e_1,\dots,e_n)} &=& K(\etrans{e_1},\dots,\etrans{e_n})\\
  \etrans{\bad} &=& \bad
\end{eqnarray*}
\caption{$\etrans{}$ -- Expression translation}
\label{etrans}
\end{figure}

@@

\begin{figure*}
\[
\begin{array}{rcl}
  \dtrans{\mbox{data } T = K_1, \dots, K_n} &=& \ktrans{\mbox{data } T = K_1, \dots, K_n} \\
  \dtrans{f \in c} &=& \ctrans{f \in c} \\
  \dtrans{f(\vec{x}) = e} &=& \forall \vec{x},y.~ \min{y} \land y =
  f(\vec{x}) \to y = \etrans{e} \\
   \raisebox{ 0.5 cm }{$\dtrans{f(x_1,\dots,x_n) = \mbox{case } e \mbox{ of } [K_i(\vec{x_i}) = e_i]}$} &\raisebox{0.5cm}{=}&  \hspace{-0.18cm} \begin{array}{rclcl}
     \forall \vec{a},y.~\mbox{min}(y) \land f(\vec{a}) = y &\rightarrow&
     \multicolumn{3}{c}{  \texttt{min}(e) \land (e = \texttt{BAD} \lor
       \bigvee_i \forall \vec{x}.~e = K_i(\vec{x}) \lor y = \texttt{UNR})} \\
     &\land& \forall \vec{x_1}.~\etrans{e} = K_1(\vec{x_1}) &\to& y = \etrans{e_1} \\
     &\land& & \cdots & \\
     &\land& \forall \vec{x_n}.~\etrans{e} = K_n(\vec{x_n}) &\to& y = \etrans{e_n}) \\
   \end{array} 
\end{array}
\]
\caption{$\dtrans{}$ -- Defintions translation}
 \label{dtrans}
\end{figure*}

@@

\begin{figure*}
 $$\ktrans{\mbox{data } T = K_1, \dots, K_n} = \Phi_1 \cup \Phi_2 \cup \Phi_3 \cup \Phi_4$$
 \hspace{5.6cm}where
  \begin{center}
    \[  \begin{array}{rcl}
      \Phi_1 &=& \bigcup_{1 \leq i \leq n}~\forall \vec{x},a.~(\min{a} \land K_i(\vec{x}) = a) \to \bigwedge_{1 \leq j \leq k} x_j = sel_{j,K_i}(a) \\
      \Phi_2 &=& \bigcup_{1 \leq i < j \leq n}~\forall \vec{x},\vec{y},a.~ \lnot (\min{a} \land K_i(\vec{x}) = a \land K_j(\vec{x}) = a) \\
      \Phi_3 &=& \bigcup_{1 \leq i \leq n}~\forall \vec{x},a.~\min{a} \land a = K_i(\vec{x}) \to ((\bigwedge_{1 \leq j \leq k} \cf{x_j}) \tot \cf{K_i(\vec{x})}) \\
      \Phi_4 &=& \bigcup_{1 \leq i \leq n}~\forall \vec{x},a.~(\min{a} \land a = K_i(\vec{x})) \to a \neq \unr \land a \neq \bad 
    \end{array} \]
  \end{center}
  \caption{$\ktrans{}$ -- Data type translation}
  \label{ktrans}
\end{figure*}

@@

\begin{figure*}
\begin{eqnarray*}
  \ctrans{e \in \{x \mid b(x) \}} &=&  \min{\etrans{b(e)}} \land (\etrans{b(e)} = true \lor \etrans{e} = \unr)\\
  \ctrans{e \in x:c_1 \to c_2(x)} &=& \forall x.~\min{\etrans{e(x)}} \to (\ctrans{x \not \in c_1} \lor  \ctrans{\etrans{e(x)} \in c_2(x)})\\
  \ctrans{e \in \cfc} &=& \cf{\etrans{e}}
\end{eqnarray*}

@@

\subsubsection{The proper way to check a module}
Consider the following situation, where $a$'s definition relies on $f$
and $g$.
\begin{center}
\includegraphics[scale=0.5]{flow.png}
\end{center}

>>>>

\def\ruleform#1{{\setlength{\fboxrule}{1.2pt}\fbox{\normalsize $#1$}}}
\newcommand{\etrans}[1]{\mathcal{E} \llbracket #1 \rrbracket}
\newcommand{\dtrans}[1]{\mathcal{D} \llbracket #1 \rrbracket}
\newcommand{\ktrans}[1]{\mathcal{K} \llbracket #1 \rrbracket}
\newcommand{\ctrans}[1]{\mathcal{C} \llbracket #1 \rrbracket}
\newcommand{\trans}[1]{\llbracket #1 \rrbracket}

@@

\begin{figure}[h]
  \centering
  \[
  \begin{array}{rclr}
    mod &:=& def_1,\dots,def_n &\\
    def &\in& \mbox{Definition} & \\
    def &:=& \texttt{data } T = K_1 \mid \cdots \mid K_n & \mbox{Data definition}\\
    &\mid& f \in c & \mbox{Contract claim}\\
    &\mid& f~\vec{x}~=~e & \\
    &\raisebox{ 0.18 cm }{$\mid$}& \multicolumn{2}{c}{
      \begin{array}{rcl}
        \hspace{-0.2cm} f~\vec{x} &=& \texttt{case } e \texttt{ of } \\
         && K_1(\vec{x}_1) \to e_1 \mid \cdots \mid K_n(\vec{x}_n) \to e_n \\
       \end{array} 
       } \\
  \end{array} \]

@@

\begin{figure}
$$
\begin{array}{rcl}
\multicolumn{3}{c}{\ruleform{\etrans{e} = t}} \\[2.5mm]
 \etrans{x} &=& x\\
  \etrans{f(e_1,\dots,e_n)} &=& f(\etrans{e_1},\dots,\etrans{e_n})\\
  \etrans{K(e_1,\dots,e_n)} &=& K(\etrans{e_1},\dots,\etrans{e_n})\\
  \etrans{\bad} &=& \bad
\end{array}
$$
\caption{$\etrans{}$ -- Expression translation}
\label{etrans}
\end{figure}

@@

\begin{figure*}
\[
\begin{array}{rcl}
  \multicolumn{3}{c}{\ruleform{\dtrans{def} = \Phi}} \\[2.5mm]
  \dtrans{\mbox{data } T = K_1, \dots, K_n} &=& \ktrans{\mbox{data } T = K_1, \dots, K_n} \\
  \dtrans{f \in c} &=& \ctrans{f \in c} \\
  \dtrans{f(\vec{x}) = e} &=& \forall \vec{x},y.~ \min{y} \land y =
  f(\vec{x}) \to y = \etrans{e} \\
   \raisebox{ 0.5 cm }{$\dtrans{f(x_1,\dots,x_n) = \mbox{case } e \mbox{ of } [K_i(\vec{x_i}) = e_i]}$} &\raisebox{0.5cm}{=}&  \hspace{-0.18cm} \begin{array}{rclcl}
     \forall \vec{a},y.~\mbox{min}(y) \land f(\vec{a}) = y &\rightarrow&
     \multicolumn{3}{c}{  \texttt{min}(e) \land (e = \texttt{BAD} \lor
       \bigvee_i \forall \vec{x}.~e = K_i(\vec{x}) \lor y = \texttt{UNR})} \\
     &\land& \forall \vec{x_1}.~\etrans{e} = K_1(\vec{x_1}) &\to& y = \etrans{e_1} \\
     &\land& & \cdots & \\
     &\land& \forall \vec{x_n}.~\etrans{e} = K_n(\vec{x_n}) &\to& y = \etrans{e_n}) \\
   \end{array} 
\end{array}
\]
\caption{$\dtrans{}$ -- Defintions translation}
 \label{dtrans}
\end{figure*}

@@

\begin{figure*}
  $$ \ruleform{\ktrans{data~def} = \Phi} $$
$$\ktrans{\mbox{data } T = K_1, \dots, K_n} = \Phi_1 \cup \Phi_2 \cup \Phi_3 \cup \Phi_4$$
 \hspace{5.6cm}where
  \begin{center}
    \[  \begin{array}{rcl}
      \Phi_1 &=& \bigcup_{1 \leq i \leq n}~\forall \vec{x},a.~(\min{a} \land K_i(\vec{x}) = a) \to \bigwedge_{1 \leq j \leq k} x_j = sel_{j,K_i}(a) \\
      \Phi_2 &=& \bigcup_{1 \leq i < j \leq n}~\forall \vec{x},\vec{y},a.~ \lnot (\min{a} \land K_i(\vec{x}) = a \land K_j(\vec{x}) = a) \\
      \Phi_3 &=& \bigcup_{1 \leq i \leq n}~\forall \vec{x},a.~\min{a} \land a = K_i(\vec{x}) \to ((\bigwedge_{1 \leq j \leq k} \cf{x_j}) \tot \cf{K_i(\vec{x})}) \\
      \Phi_4 &=& \bigcup_{1 \leq i \leq n}~\forall \vec{x},a.~(\min{a} \land a = K_i(\vec{x})) \to a \neq \unr \land a \neq \bad 
    \end{array} \]
  \end{center}
  \caption{$\ktrans{}$ -- Data type translation}
  \label{ktrans}
\end{figure*}

@@

\begin{figure*}
  $$ \ruleform{\ctrans{contract} = \Phi} $$
\begin{eqnarray*}
  \ctrans{e \in \{x \mid b(x) \}} &=&  \min{\etrans{b(e)}} \land (\etrans{b(e)} = true \lor \etrans{e} = \unr)\\
  \ctrans{e \in x:c_1 \to c_2(x)} &=& \forall x.~\min{\etrans{e(x)}} \to (\ctrans{x \not \in c_1} \lor  \ctrans{\etrans{e(x)} \in c_2(x)})\\
  \ctrans{e \in \cfc} &=& \cf{\etrans{e}}
\end{eqnarray*}

@@

\subsubsection{The proper way to check a module}
Consider the following situation, where $a$'s definition relies on $f$
and $g$.
\begin{center}
\includegraphics[scale=0.3]{flow.png}
\end{center}
